Making ice sheet models \\ scale properly

### 2023-01-06 talk/slides.tex
more progress I hope

\begin{frame}{simplified ice sheet model (strong form)}

@@

@@

\begin{frame}{ice thickness and ``mass continuity equation'' view}

\begin{itemize}
\item a helpful viewpoint on this complicated system is the equation

FIXME

where $\bU$ is vertically-averaged horizontal velocity within icy domain $\Lambda(t)$
\item \emph{question:} is it really an advection?

FIXME

\item[] \emph{answer:} no, because ice flows (mostly) downhill

\item alternative equation, the \emph{shallow diffusion} model of ice sheets:

FIXME
\bigskip
\item<2> {\footnotesize for more on the shallow diffusion perspective see Jouvet \& Bueler (2012)}
\end{itemize}
\end{frame}

>>>>

\begin{frame}{basic ice sheet model (strong form)}

@@

\begin{frame}{the mass-continuity equation view}

\begin{itemize}
\item a thickness-transport form exposes the basic ice sheet model
\item define:
\begin{align*}
H(t,x,y) &=s-b && \text{ice thickness} \\
\bU(t,x,y) &= \frac{1}{H} \int_b^s \bu \,dz && \begin{matrix} \text{vertically-averaged} \\ \hspace{-1.8mm} \text{horizontal velocity} \end{matrix}
\end{align*}
\item the \alert{mass continuity equation} follows from the surface kinematical equation and incompressibility:
   $$\frac{\partial H}{\partial t} + \nabla \cdot \left(\bU H\right) = a$$
\item \emph{question:} is it really an advection?
\item[] \emph{answer:} not really, because ice flows (mostly) downhill
\item an alternative form is the \emph{shallow diffusion} equation for ice sheets:
   $$\frac{\partial H}{\partial t} = \nabla \cdot \left(D \grad s\right) + a$$
\item<2> {\footnotesize for more on the shallow diffusion perspective see Jouvet \& Bueler (2012)}
\end{itemize}
\end{frame}

\begin{frame}{shallow ice sheet models}

\begin{itemize}
\item the model on the last slide is \emph{not yet} in common use for ice sheet simulations
    \begin{itemize}
    \item[$\circ$] for now, it is too expensive at scale
    \item[$\circ$] one goal: \alert{make it fast enough to use!}
    \end{itemize}
\item<2-> current-technology models use various shallow simplifications
\item<2-> simplest is a lubrication approximation, the \alert{shallow ice approximation}, in which $- \bu|_s \cdot \bn_s$ resolves to a nonlinear differential operator  
\end{itemize}
\end{frame}

\begin{frame}{shallow ice approximation: details}

{\small
\begin{align*}
s - b &\ge 0 \\
\frac{\partial s}{\partial t} + \Phi(s) - a &\ge 0 \\
(s - b) \left(\frac{\partial s}{\partial t} + \Phi(s) - a\right) &= 0
\end{align*}
where
\begin{align*}
\Phi(s) &= - \bu|_s \cdot \bn_s \\
        &= - \frac{\gamma}{\pp} (s-b)^{\pp} |\grad s|^{\pp} - \grad \cdot\left(\frac{\gamma}{\pp+1} (s-b)^{\pp+1} |\grad s|^{\pp-2} \grad s\right)
\end{align*}

\begin{itemize}
\item $\nabla$ only acts in horizontal variables
\item constants $\pp = \nn+1$ and $\gamma > 0$ relate to ice deformation
\item $\Phi$ is a \alert{doubly-nonlinear differential operator}, simultaneously of porous medium and $\pp$-Laplacian type
\end{itemize}
}
\end{frame}

@@
